---
title: "Getting Started in ihr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started in ihr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ihr)
```



# Example Datasets
The ihr package includes two datasets: example_heart_1, and example_heart_5. The one subject data is simply the first subject of the five subject data. 

## Example data for 1 subject

Example data with 1 subject can be loaded with:

```{r}
data("example_data_1")
```
  
This dataset contains 154104 observations of 3 columns corresponding to the three components listed in the introduction:

* `"id"` - Factor (character string) column for subject identification
* `"time"` - POSIXct column for datetime values
* `"hr"` - Numeric column for heart rate
Data used with ihr functions may have additional columns, but the columns for id, time and hr values must be named as above.

```{r}
dim(example_heart_1)
str(example_heart_1)
head(example_heart_1)
```


# Vizualizations

The ihr package supports two plot types described below:

| Function call | Visualization description | 
|:------------|:------------------------------|
|`plot_hr` | Time Series plot of Heart Rate | 
|`mahe`| MAHE plot displying peak and nadirs| 

## Example plot: Time series

## Examlpe plot: MAHE

# 
